A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal

### 2020-07-16 Manuscript/PNAS-Adams-rticle.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers

\author{}
\date{\vspace{-2.5em}}

@@

\textbf{Classification:} Biological Sciences -- Evolution \hfill\break

\begin{center}
\textbf{A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal}
\end{center}

\hfill\break

\begin{center}
\textbf{Dean C. Adams$^{1 *}$, Erica K. Baken$^{1,2}$,  and Michael L. Collyer$^{2}$}
\end{center}

\hfill\break

\(^{1}\)Department of Ecology, Evolution, and Organismal Biology, Iowa
State University, Ames, Iowa, USA.

\(^{2}\)Department of Science, Chatham University, Pittsburgh,
Pennsylvania, USA.

\(^{*}\)Correspondence: Dean C. Adams. Department of Ecology, Evolution,
and Organismal Biology, Iowa State University, Ames, Iowa, USA.
515-294-3834.
\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}

\hfill\break

\begin{center}15 July, 2020\end{center}

\hfill\break

\begin{longtable}[]{@{}rr@{}}
\toprule
words & characters\tabularnewline
\midrule
\endhead
2553 & 16162\tabularnewline
\bottomrule
\end{longtable}

\newpage

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, and wish to compare the strength of that signal
across traits. However, analytical tools for such comparisons have
largely remained underdeveloped. Here we evaluate the efficacy of one
commonly used parameter (Pagel's \(\lambda\)) to estimate the strength
of phylogenetic signal in phenotypic traits, and evaluate the degree to
which \(\lambda\) correctly identifies known levels of phylogenetic
signal. We find that \(\lambda\) behaves as a Bernoulli random variable,
and that estimates are increasingly skewed at larger and smaller input
levels of phylogenetic signal. Further, the precision of \(\lambda\) in
estimating actual levels of phylogenetic signal is often inaccurate, and
biological interpretations of the strength of phylogenetic signal based
on \(\lambda\) are therefore compromised. We propose a standardized
effect size based on \(\kappa\), (\(Z_\kappa\)), which measures the
strength of phylogenetic signal more reliably than does \(\lambda\), and
places that signal on a common scale for statistical comparison. We
develop tests based on \(Z_\kappa\) to provide a mechanism for formally
comparing the strength of phylogenetic signal across datasets, in much
the same manner as effect sizes may be used to summarize patterns in
quantitative meta-analysis. Our approach extends the phylogenetic
comparative toolkit to address hypotheses that compare the strength of
phylogenetic signal between various phenotypic traits, even when those
traits are found in different evolutionary lineages or have different
units or scales. \hfill\break

\textbf{Keywords}: phylogenetic signal, macroevolution, lambda, kappa

\newpage

\hypertarget{significance-statement}{%
\section{Significance Statement}\label{significance-statement}}

Evolutionary biologists wish to quantify and compare the strength of
phylogenetic signal across traits, but analytical tools for these
comparisons are generally lacking. Here we develop a standardized effect
size based on \(\kappa\) (\(Z_\kappa\)), which measures the strength of
phylogenetic signal on a common statistical scale, and provides a
mechanism for formally comparing the strength of phylogenetic signal
across datasets. Additionally, we find that a commonly used parameter
(Pagel's \(\lambda\)) is insuitable for this purpose. Our procedure
enables biologists to quantitatively address hypotheses that compare the
strength of phylogenetic signal between various phenotypic traits, even
when those traits are found in different evolutionary lineages or have
different units or scales.

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry among species
violates the assumption of independence among trait values that is
common for statistical tests (1, 2). Accounting for this evolutionary
non-independence is the purview of \emph{phylogenetic comparative
methods} (PCMs): a suite of analytical tools that condition trends in
the data on the phylogenetic relatedness of observations (3--12). These
methods are predicated on the notion that phylogenetic signal -- the
tendancy for closely related species to display similar trait values --
is present in cross-species datasets (1, 13, 14). Indeed, under numerous
evolutionary models, phylogenetic signal is to be expected, as
stochastic character change along the hierarchical structure of the tree
of life generates trait covaration among related taxa (1, 14, 15).
\hfill\break

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets (13, 14, 16--19), and their statistical
properties -- namely type I error rates and statistical power -- have
been investigated to determine under what conditions phylogenetic signal
can be detected (15, 18, 20--25). One of the most widely used methods
for characterizing phylogenetic signal is Pagel's \(\lambda\) (13),
which transforms the lengths of the internal branches of the phylogeny
to improve the fit of data to the phylogeny via maximum likelihood (13,
26). When incorporated in PGLS, \(\lambda\) serves as a tuning parameter
(ranging from \(0\rightarrow1\)) which is optimized via log-likelihood
profiling while evaluating the covariation between the dependent and
independent variables, given the phylogeny (13, 26). To infer whether
phylogenetic signal differs from no signal or a Brownian motion model of
evolutionary divergence, the observed model fit using \(\hat\lambda\)
may be statistically compared to that using \(\lambda=0\) or
\(\lambda=1\) via likelihood ratio tests (26--28) or confidence limits
(29). Another widely used measure is Blomberg's \(\kappa\) (14), which
measures the ratio of observed trait variation to the amount expected
under Brownian motion. Blomberg's \(\kappa\) can be treated as a test
statistic by employing a permutation test to generate its sampling
distribution (14, 18) for determining whether significant phylogenetic
signal is present in data. Both \(\lambda\) and \(\kappa\) seem
intuitive to interpret, as a value of \(0\) for both corresponds to no
phylogenetic signal, while a value of \(1\) corresponds to the amount of
phylogenetic signal expected under Brownian motion. Thus, it is tempting
to regard both \(\lambda\) and \(\kappa\) as descriptive statistics that
measure the relative strength of phylogenetic signal, providing an
estimate of its magnitude for comparison. \hfill\break

The appeal of Pagel's \(\lambda\) and Blomberg's \(\kappa\) as
descriptive statistics is that they provide a basis for interpreting
``weak'' versus ``strong'' phylogenetic signal; i.e., small versus large
values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
comparative sense (30--32). Nonetheless, an important question that has
yet to be considered is whether these statistics are or can be converted
to effect sizes for comparative analyses across datasets? To be
statistics representing phylogenetic signal, they should have reliable
distributional properties, which could be revealed with simulation
experiments. For instance, as a proportional random variable bounded by
\(0\) and \(1\), we might expect that \(\hat{\lambda}\) follows a
Bernoulli distribution (\textbf{add ref}); i.e., branch lengths in a
tree are scaled proportionally to the probability that data arise from a
BM process. Given a known \(\lambda\) value used to generate random data
on a tree, we would also expect that the mean of an empirical sampling
distribution of \(\hat{\lambda}\) would approximately equal \(\lambda\);
the dispersion of \(\hat{\lambda}\) would be largest at intermediate
values of \(\lambda\), \(\hat{\lambda}\) would be predictable over the
range of \(\lambda\) with respect to treesize; the distribution of
\(\hat{\lambda}\) would be symmetric at intermediate values of
\(\lambda\) and more skewed toward values of 0 or 1; and that the
distribution of \(\hat{\lambda}\) will be more platykurtic at
intermediate values of \(\lambda\), becoming more leptokurtic toward 0
and 1 (\textbf{add same ref}). Prior work (20) seems to support some of
these conjectures, based superficially on statistical moments for a
given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of
(20)). However, because the ``strength of Brownian motion'' was
simulated as a varied weighted-average of data simulated on trees with
\(\lambda=0\) and \(\lambda=1\) and not prescribed values of \(\lambda\)
(20), interpretation of these patterns is challenging. \hfill\break

By contrast, Blomberg's \(\kappa\), which is positively ubounded, might
be expected to follow a normal distribution (\textbf{add ref}). Thus we
might expect that \(\kappa\) is symmetrically distributed across
different strengths of phylogenetic signal, for any \(\lambda\) used to
generate data. This attribute seemed less reasonable based on the
simulations performed by Münkemüller et al.~(20), as distibutions were
postively skewed, suggesting the Blomberg's \(\kappa\) might not behave
as a statistic that follows a noraml distribution. However, because
their simulations used a weighted combination of simulated phylogenetic
signal strengths, strong inferences are not possible (and distributional
attributes were not the intended result of their simulations). For both
Pagel's \(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical
moments across a range of \(\lambda\) used to generate data would be
valuable for adjudicating the relibalilty of these statistics.
Furthermore, these are statistics that appear to have expected values
that vary with tree size (20), making comparisons across studies
challenging. Therefore, transformation of these statistics into
\(Z\)-scores in the same simulation experiments would allow evaluation
of the efficacy of each statistic to yield effect sizes that could be
used for comparative anlyses. \hfill\break

@@

\textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
innacurate.} Computer simulations revealed that for \(\hat{\lambda}\),
the distributional expectations of a Bernoulli variable were mostly
upheld. First, the mean value of \(\hat{\lambda}\) increased as
\(\lambda\) increased (Fig. 1 black line); the standard deviation of
\(\hat{\lambda}\) was largest at intermediate values of \(\lambda\) and
smallest at extreme values (Fig. 1 red line); and the distributions of
\(\hat{\lambda}\) were increasing skewed at extreme values of
\(\lambda\) but rather normal at intermediate values (Fig. 1 blue line).
For small tree sizes, it was also clear that distributions were more
platykurtik at intermeidate values of \(\hat{\lambda}\). Second,
\(\hat{\lambda}\) was a negatively-biased estimate of \(\lambda\) for
small tree sizes, as the mean value was consistently less than the input
\(\lambda\) value. Third, standard deviations of \(\hat{\lambda}\) were
negatively associated with tree size, and for tress of 128 species or
less, \(\hat{\lambda}\) was quite variable, except for cases when
\(\lambda\) was near or equal to \(1\). As such its statistical
properties make it insuitable as an effect size for measuring the
strength of phylogenetic signal. \hfill\break

\textbf{Kappa (\(\kappa\)) estimates of phylogenetic signal are more
stable.} Results for \(\kappa\) demonstrated better statistical
properties. First, mean values increased consistently with \(\lambda\)
irespective of tree size (Fig. 2 black line); the standard deviation of
\(\kappa\) was consistent across tree sizes (Fig. 2 red line); and
\(\kappa\) was normally distributed regardless of \(\lambda\) or tree
size (Fig. 2 blue line). The standard deviation of \(\kappa\) increased
with \(\lambda\), somewhat tracking the mean, but was always less than
the corresponding means. This is perhaps unsurprising, as \(\kappa\) is
bound by 0, and was never large for small values of \(\lambda\).
Finally, distributions of \(\kappa\) displayed slight skewing only at
small tree sizes and when \(\lambda\) was near or equal to 1. This
differs from the results of (20), where skewing appears to be from
combining random values generated independently. \hfill\break

\textbf{Effect sizes from \(\kappa\) (\(Z_{\kappa}\)) better
characterize phylogenetic signal.} We derived effect sizes (Z-scores)
for both \(\lambda\) and \(\kappa\) (see Methods) and evaluated them
relative to known phylogenetic signal. Both \(Z_{\lambda}\) and
\(Z_{\kappa}\) were associated with input phylogenetic signal
(\(\lambda\)), indicating that both statistics captured the observed
signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
little sense, as they were more strongly associated with tree size than
the actual phylogenetic signal in the data (Fig. 3). By contrast,
\(Z_{\kappa}\) was much more consistent accross tree sizes, and
exhibited a much stronger association with phylogenetic signal strength
as compared to tree size (Fig. 3). Thus between the two statistics,
\(Z_{\kappa}\) is a more reliable measure of the strength of
phylogenetic signal, and may be used to compare levels of phylogenetic
signal across datasets. \hfill\break

@@

@@

address hypotheses that compare the strength of phylogenetic signal
between various phenotypic traits, even when those traits are found in
different evolutionary lineages or have different units or scales

>>>>

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\templatetype{pnasresearcharticle}  % Choose template

\title{A Standardized Effect Size for Evaluating and Comparing the Strength of
Phylogenetic Signal}

\author[a]{Dean C. Adams}
\author[a,b]{Erica K. Baken}
\author[b]{Michael L. Collyer}

  \affil[a]{Department of Ecology, Evolution, and Organismal Biology, Iowa State
University, Ames, Iowa, USA.}
  \affil[b]{Department of Science, Chatham University, Pittsburgh, Pennsylvania,
USA.}

% Please give the surname of the lead author for the running footer
\leadauthor{Adams et al.}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Evolutionary biologists wish to quantify and compare the strength of
phylogenetic signal across traits, but analytical tools for these
comparisons are generally lacking. Here we develop a standardized effect
size based on \(\kappa\) (\(Z_\kappa\)), which measures the strength of
phylogenetic signal on a common statistical scale, and provides a
mechanism for formally comparing the strength of phylogenetic signal
across datasets. Additionally, we find that a commonly used parameter
(Pagel's \(\lambda\)) is insuitable for this purpose. Our procedure
enables biologists to quantitatively address hypotheses that compare the
strength of phylogenetic signal between various phenotypic traits, even
when those traits are found in different evolutionary lineages or have
different units or scales.}

\authorcontributions{D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the
research and wrote the paper.}

\authordeclaration{The authors declare no conflict of interest.}

\correspondingauthor{\textsuperscript{} }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  phylogenetic signal |  macroevolution |  lambda |  kappa  }

\begin{abstract}
Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, and wish to compare the strength of that signal
across traits. However, analytical tools for such comparisons have
largely remained underdeveloped. Here we evaluate the efficacy of one
commonly used parameter (Pagel's \(\lambda\)) to estimate the strength
of phylogenetic signal in phenotypic traits, and evaluate the degree to
which \(\lambda\) correctly identifies known levels of phylogenetic
signal. We find that \(\lambda\) behaves as a Bernoulli random variable,
and that estimates are increasingly skewed at larger and smaller input
levels of phylogenetic signal. Further, the precision of \(\lambda\) in
estimating actual levels of phylogenetic signal is often inaccurate, and
biological interpretations of the strength of phylogenetic signal based
on \(\lambda\) are therefore compromised. We propose a standardized
effect size based on \(\kappa\), (\(Z_\kappa\)), which measures the
strength of phylogenetic signal more reliably than does \(\lambda\), and
places that signal on a common scale for statistical comparison. We
develop tests based on \(Z_\kappa\) to provide a mechanism for formally
comparing the strength of phylogenetic signal across datasets, in much
the same manner as effect sizes may be used to summarize patterns in
quantitative meta-analysis. Our approach extends the phylogenetic
comparative toolkit to address hypotheses that compare the strength of
phylogenetic signal between various phenotypic traits, even when those
traits are found in different evolutionary lineages or have different
units or scales.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

@@

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{We thank E. Glynne and B. Juarez for comments on early drafts of the
manuscript. This work was supported in part by NSF grant DBI-1902511 (to
D.C.A.) and DBI-1902694 (to M.L.C.). \hfill\break}

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry among species
violates the assumption of independence among trait values that is
common for statistical tests (1, 2). Accounting for this evolutionary
non-independence is the purview of \emph{phylogenetic comparative
methods} (PCMs): a suite of analytical tools that condition trends in
the data on the phylogenetic relatedness of observations (3--12). These
methods are predicated on the notion that phylogenetic signal -- the
tendancy for closely related species to display similar trait values --
is present in cross-species datasets (1, 13, 14). Indeed, under numerous
evolutionary models, phylogenetic signal is to be expected, as
stochastic character change along the hierarchical structure of the tree
of life generates trait covaration among related taxa (1, 14, 15).

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets (13, 14, 16--19), and their statistical
properties -- namely type I error rates and statistical power -- have
been investigated to determine under what conditions phylogenetic signal
can be detected (15, 18, 20--25). One of the most widely used methods
for characterizing phylogenetic signal is Pagel's \(\lambda\) (13),
which transforms the lengths of the internal branches of the phylogeny
to improve the fit of data to the phylogeny via maximum likelihood (13,
26). When incorporated in PGLS, \(\lambda\) serves as a tuning parameter
(ranging from \(0\rightarrow1\)) which is optimized via log-likelihood
profiling while evaluating the covariation between the dependent and
independent variables, given the phylogeny (13, 26). To infer whether
phylogenetic signal differs from no signal or a Brownian motion model of
evolutionary divergence, the observed model fit using \(\hat\lambda\)
may be statistically compared to that using \(\lambda=0\) or
\(\lambda=1\) via likelihood ratio tests (26--28) or confidence limits
(29). Another widely used measure is Blomberg's \(\kappa\) (14), which
measures the ratio of observed trait variation to the amount expected
under Brownian motion. Blomberg's \(\kappa\) can be treated as a test
statistic by employing a permutation test to generate its sampling
distribution (14, 18) for determining whether significant phylogenetic
signal is present in data. Both \(\lambda\) and \(\kappa\) seem
intuitive to interpret, as a value of \(0\) for both corresponds to no
phylogenetic signal, while a value of \(1\) corresponds to the amount of
phylogenetic signal expected under Brownian motion. Thus, it is tempting
to regard both \(\lambda\) and \(\kappa\) as descriptive statistics that
measure the relative strength of phylogenetic signal, providing an
estimate of its magnitude for comparison.

The appeal of Pagel's \(\lambda\) and Blomberg's \(\kappa\) as
descriptive statistics is that they provide a basis for interpreting
``weak'' versus ``strong'' phylogenetic signal; i.e., small versus large
values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
comparative sense (30--32). Nonetheless, an important question that has
yet to be considered is whether these statistics are or can be converted
to effect sizes for comparative analyses across datasets? To be
statistics representing phylogenetic signal, they should have reliable
distributional properties, which could be revealed with simulation
experiments. For instance, as a proportional random variable bounded by
\(0\) and \(1\), we might expect that \(\hat{\lambda}\) follows a
Bernoulli distribution (\textbf{add ref}); i.e., branch lengths in a
tree are scaled proportionally to the probability that data arise from a
BM process. Given a known \(\lambda\) value used to generate random data
on a tree, we would also expect that the mean of an empirical sampling
distribution of \(\hat{\lambda}\) would approximately equal \(\lambda\);
the dispersion of \(\hat{\lambda}\) would be largest at intermediate
values of \(\lambda\), \(\hat{\lambda}\) would be predictable over the
range of \(\lambda\) with respect to treesize; the distribution of
\(\hat{\lambda}\) would be symmetric at intermediate values of
\(\lambda\) and more skewed toward values of 0 or 1; and that the
distribution of \(\hat{\lambda}\) will be more platykurtic at
intermediate values of \(\lambda\), becoming more leptokurtic toward 0
and 1 (\textbf{add same ref}). Prior work (20) seems to support some of
these conjectures, based superficially on statistical moments for a
given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of
(20)). However, because the ``strength of Brownian motion'' was
simulated as a varied weighted-average of data simulated on trees with
\(\lambda=0\) and \(\lambda=1\) and not prescribed values of \(\lambda\)
(20), interpretation of these patterns is challenging.

By contrast, Blomberg's \(\kappa\), which is positively ubounded, might
be expected to follow a normal distribution (\textbf{add ref}). Thus we
might expect that \(\kappa\) is symmetrically distributed across
different strengths of phylogenetic signal, for any \(\lambda\) used to
generate data. This attribute seemed less reasonable based on the
simulations performed by Münkemüller et al.~(20), as distibutions were
postively skewed, suggesting the Blomberg's \(\kappa\) might not behave
as a statistic that follows a noraml distribution. However, because
their simulations used a weighted combination of simulated phylogenetic
signal strengths, strong inferences are not possible (and distributional
attributes were not the intended result of their simulations). For both
Pagel's \(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical
moments across a range of \(\lambda\) used to generate data would be
valuable for adjudicating the relibalilty of these statistics.
Furthermore, these are statistics that appear to have expected values
that vary with tree size (20), making comparisons across studies
challenging. Therefore, transformation of these statistics into
\(Z\)-scores in the same simulation experiments would allow evaluation
of the efficacy of each statistic to yield effect sizes that could be
used for comparative anlyses.

@@

\textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
innacurate.} Computer simulations revealed that for \(\hat{\lambda}\),
the distributional expectations of a Bernoulli variable were mostly
upheld. First, the mean value of \(\hat{\lambda}\) increased as
\(\lambda\) increased (Fig. 1 black line); the standard deviation of
\(\hat{\lambda}\) was largest at intermediate values of \(\lambda\) and
smallest at extreme values (Fig. 1 red line); and the distributions of
\(\hat{\lambda}\) were increasing skewed at extreme values of
\(\lambda\) but rather normal at intermediate values (Fig. 1 blue line).
For small tree sizes, it was also clear that distributions were more
platykurtik at intermeidate values of \(\hat{\lambda}\). Second,
\(\hat{\lambda}\) was a negatively-biased estimate of \(\lambda\) for
small tree sizes, as the mean value was consistently less than the input
\(\lambda\) value. Third, standard deviations of \(\hat{\lambda}\) were
negatively associated with tree size, and for tress of 128 species or
less, \(\hat{\lambda}\) was quite variable, except for cases when
\(\lambda\) was near or equal to \(1\). As such its statistical
properties make it insuitable as an effect size for measuring the
strength of phylogenetic signal.

\begin{figure}
\centering
\includegraphics{new.fig.1.temp.png}
\caption{Patterns in \(\lambda\).{}}
\end{figure}

\textbf{Kappa (\(\kappa\)) estimates of phylogenetic signal are more
stable.} Results for \(\kappa\) demonstrated better statistical
properties. First, mean values increased consistently with \(\lambda\)
irespective of tree size (Fig. 2 black line); the standard deviation of
\(\kappa\) was consistent across tree sizes (Fig. 2 red line); and
\(\kappa\) was normally distributed regardless of \(\lambda\) or tree
size (Fig. 2 blue line). The standard deviation of \(\kappa\) increased
with \(\lambda\), somewhat tracking the mean, but was always less than
the corresponding means. This is perhaps unsurprising, as \(\kappa\) is
bound by 0, and was never large for small values of \(\lambda\).
Finally, distributions of \(\kappa\) displayed slight skewing only at
small tree sizes and when \(\lambda\) was near or equal to 1. This
differs from the results of (20), where skewing appears to be from
combining random values generated independently.

\begin{figure}
\centering
\includegraphics{new.fig.2.temp.png}
\caption{Patterns in \(\kappa\).{}}
\end{figure}

\textbf{Effect sizes from \(\kappa\) (\(Z_{\kappa}\)) better
characterize phylogenetic signal.} We derived effect sizes (Z-scores)
for both \(\lambda\) and \(\kappa\) (see Methods) and evaluated them
relative to known phylogenetic signal. Both \(Z_{\lambda}\) and
\(Z_{\kappa}\) were associated with input phylogenetic signal
(\(\lambda\)), indicating that both statistics captured the observed
signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
little sense, as they were more strongly associated with tree size than
the actual phylogenetic signal in the data (Fig. 3). By contrast,
\(Z_{\kappa}\) was much more consistent accross tree sizes, and
exhibited a much stronger association with phylogenetic signal strength
as compared to tree size (Fig. 3). Thus between the two statistics,
\(Z_{\kappa}\) is a more reliable measure of the strength of
phylogenetic signal, and may be used to compare levels of phylogenetic
signal across datasets.

\begin{figure}
\centering
\includegraphics{new.fig.3.temp.png}
\caption{Patterns in Z.{}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{new.fig.4.png}
\caption{(A) Linear measures for relative body size, and regions of the
body used to estimate surface area to volume (SA:V) ratios. (B)
Permutation distributions of phylogenetic signal for SA:V and
\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\), with their
95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
\end{figure}

@@

\textbf{To be edited.} address hypotheses that compare the strength of
phylogenetic signal between various phenotypic traits, even when those
traits are found in different evolutionary lineages or have different
units or scales.

It is common in comparative evolutionary studies to characterize the
phylogenetic signal in phenotypic traits to determine the extent to
which shared evolutionary history has generated trait covariation among
taxa. However, while numerous analytical approaches may be used to
quantify phylogenetic signal (13, 14, 16--18), methods that explicitly
measure the strength of phylogenetic signal, or facilitate comparisons
among datasets, have remained underdeveloped. In this study, we
evaluated the precision of one common measure, Pagel's \(\lambda\), and
explored its efficacy for characterizing the strength of phylogenetic
signal in phenotypic data. Using computer simulations, we found that the
precision of \(\lambda\) increased with increasing sample sizes; a
pattern noted previously (25), and one that conformed with parametric
statistical theory (35). However, we also found that vastly different
\(\lambda\) estimates could be obtained from data containing the same
level of phylogenetic signal, and that similar \(\lambda\) estimates may
be obtained from data containing differing levels of phylogenetic
signal. Further, the precision of \(\lambda\) varied with the strength
of phylogenetic signal, where lower precision was observed when in data
whose phylogenetic signal was of intemediate strength. From these
findings we conclude that \(\lambda\) is not a reliable indicator of the
observed strength of phylogenetic signal in phenoytpic datasets, and
that biological interpretations of the strength of signal based on this
parameter may innacurately characterize such effects.

As an alternative, we described a standardized effect size (\(Z\)) for
assessing the strength of phylogenetic signal. \(Z\) expresses the
magnitude of phylogenetic signal as a standard normal deviate, which is
easily interpretable as the strength of phylogenetic signal relative to
the mean. We applied this concept to both \(\lambda\) and \(\kappa\),
and found that \(Z_\kappa\) was a better estimate of the strength of
phylogenetic signal in phenotypic data. First, \(Z_\kappa\) was more
precise than \(Z_\lambda\), and precision was more consistent across the
range of input levels of phylogenetic signal. Additionally, values of
\(Z_\kappa\) more accurately tracked known changes in the magnitude of
phylogenetic signal, as demonstrated by the linear relationship between
\(Z_\kappa\) and \(\lambda_{in}\). Thus, \(Z_\kappa\) holds promise as a
measure of the relative strength of phylogenetic signal that reflects
the magnitude of this effect in phenotypic data. We therefore recommend
that future studies interested in the strength of phylogenetic signal
incorporate \(Z_\kappa\) as a statistical measure of this effect.

Based on the effect size \(Z_\kappa\), we then proposed a two-sample
test, which provides means of determining whether the strength of
phylogenetic signal is greater in one phenotypic trait as compared to
another, via a hypothesis test. Prior studies have summarized patterns
of variation in phylogenetic signal across datasets using summary test
values, such as \(\kappa\) (14). However, \(\kappa\) does not scale
linearly with input levels of phylogenetic signal, and its variance
increases (i.e., precision decreases) with increasing strength of
phylogenetic signal (20, 22). Thus, \(\kappa\) should not be considered
an effect size that measures the strength of phylogenetic signal on a
common scale. By contrast, standardizing \(\kappa\) (\(Z_\kappa\), via
equation 2) alleviates these concerns, and facilitates formal
statistical comparisons of the strength of signal across datasets. Thus
when viewed from this perspective, the approach developed here aligns
well with other statistical approaches such as meta-analysis (36--38),
where summary statistics across datasets are converted to standardized
effect sizes for subsequent ``higher order'' statistical summaries or
comparisons. As such, our approach enables evolutionary biologists to
quantitatively examine the relative strength of phylogenetic signal
across a wide range of phenotypic traits, and thus opens the door for
future discoveries that inform on how phenotypic diversity accumulates
in macroevolutionary time across the tree of life.

One important advantage of the approach advocated here is that the
resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
measurement cancel out during the calculation of \(Z\) (39). Thus,
\(Z_\kappa\) represents the strength of phylogenetic signal on a common
and comparable scale -- measured in standard deviations -- regardless of
the initial units and original scale of the phenotypic variables under
investigation. This means that the strength of phylogenetic signal may
be compared across datsets for continuous phenotypic traits measured in
different units and scale, because those units have been standardized
through their conversion to \(Z_\kappa\). For example, our approach
could be utilized to determine whether the strength of phylogenetic
signal (say, in response to ecological differentiation) is stronger in
morphological traits (linear traits: \(mm\)), physiological traits
(metabolic rate: \(\frac{O^2}{min}\)), or behavioral traits (aggression:
\(\frac{\#{displays}}{second}\)). In fact, our empirical example
provided such a comparison, as SA:V is represented in \(mm^{-1}\) while
relative body size is a unitless ratio (\(\frac{BW}{SVL}\)).
Additionally, our method is capable of comparing the strength of
phylogenetic signal in traits of different dimensionality, as estimates
of phylogenetic signal using \(\kappa\) have been generalized for
multivariate data (18). Furthermore, tests based on \(\hat{Z}_{12}\) may
be utilized for comparing the strength of phylogenetic signal among
datasets containing a different number of species, and even for
phenotypes obtained from species in different lineages, because their
phylogenetic non-independence and observed variation are taken into
account in the generation of the empirical sampling distribution via
permutation.
